Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-08 cluster_paper/ms.tex
update ms

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{Core \& Halo of NGC\,2516}

@@

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  %An Open Cluster With a 300\,pc Halo
  An Open Cluster Spanning 500\,Parsecs
  %NGC\,2516  300\,pc
}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  It is
  important to verify that these ``halos'', ``tails'', and ``strings''
  are of similar ages and compositions as stars in the denser part of
  the cluster.  We present an analysis of NGC\,2516
  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
  and an apparent halo of stars spanning 500\,pc.  Combining
  photometry from Gaia, rotation periods from TESS, and lithium
  measurements from Gaia-ESO and GALAH, we find that the halo of
  NGC\,2516 is real and coeval with the cluster's core.  Two thirds of
  kinematically selected halo members out to 250\,pc from the cluster
  center have rotation periods consistent with a gyrochronological age
  of 150\,Myr.  A comparison sample of field stars shows no such
  trend.  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of
  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
  paradigm in which the halos of many open clusters are more populous
  than their cores.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following intervals of right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample.%, which led to
%the omission of 1.1\% of the stars drawn from the volume noted above.

@@

% Are these halo stars truly associated with the cluster's core?
% What is the 
% The opportunity of verifying and understanding the diffuse
% population of \cn\ prompted our interest in understanding
% the reality and origin of the halo.

%Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
%$\mu_\delta$ projection were reported almost entirely by
%\citet{meingast_2021}.

An initial step in visualizing the reported structure and kinematics
of the candidate cluster members is to consider them in Cartesian
coordinates (Figure~\ref{fig:XYZ}).  We computed the physical
positions by transforming from $(\alpha, \delta, \pi)$ to
galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}.  The direction of galactic rotation is
$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
median parameters for core members for which
\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.

@@

A second possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 300\,pc
(Figure~\ref{fig:gaia6d}).  Some empirical evidence for differential
reddening is discussed in Appendix~\ref{app:diffred}. Based on the
available evidence, we expect that both field star contamination and
differential reddening could play a role in the larger scatter of the
halo relative to the core.  A third possibility that we have not
explored is whether the binary fraction could also differ between the
different regions of the cluster.

@@

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  %Their statistical uncertainties ($\approx
%25$\,Myr) were comparable to the systematic uncertainties.

Our goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for the more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt a MSTO age for \cn\ of $175\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of the statistical
and systematic components from \cite{cummings_2018}. 
%The relative age of ``slightly older than
%the Pleiades'' is likely the more robust statement, consistent with
%reports since \citet{mermilliod_comparative_1981}.

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.
The ``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrp<1.2$.  The implications of this comparison for the age
of \cn\ are discussed in Section~\ref{disc:absage}.
%An important note is that stars redder than $\bpmrp\gtrsim 2.2$
%(later than $\approx$M2V)  are beyond the magnitude limit of our survey.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members that meet kinematic
  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
  rotation periods are shown in blue;  halo members for which rotation
  periods should have been detectable, but were not detected, are
  shown in orange.  Stars in the core are shown as black points.
  Computing the 2D-tangential velocity requires a correction for
  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
  right}, the histograms on bin versus 3D separation from the core and
  2D tangential velocity difference.  Bin widths are chosen to enforce
  the same number of stars in the denominator ($N=54$) per bin.  In the
  {\it bottom panels}, the innermost and outermost 100 stars in
  position and velocity space are displayed in the space of rotation
  periods and dereddened colors.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
			\includegraphics[width=1\textwidth]{f5a.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with Gaia-ESO
  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Field stars (gray
  points) are Keck-HIRES measurements of stars with detected
  transiting planets from the Kepler mission
  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in violet
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
		\includegraphics[width=1\textwidth]{f5b.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
  GALAH surveys, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ (including
  non-detections) are shown with gray crosses.  The lithium-rotation
  correlation for the K-dwarfs has been observed in other young
  clusters, and is discussed in Section~\ref{subsec:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

@@

%
%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
%
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.  We then measured the lithium equivalent widths using
the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
by first focusing in on a 15\AA\ window centered on the 6707.835\AA
lithium doublet.  We then continuum normalized the spectra using a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
to the continuum normalized spectra, considering only a $\pm 1\,$\AA
window centered on the Li doublet.  The equivalent widths were then
evaluated by numerically integrating the fitted model over the same
window.  Our approach therefore includes the Fe\,6707.44 blend in the
reported Li equivalent widths---which leads to systematic
overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
uncertainties on the EWs, we repeated the procedure twenty times, but
added noise to the spectra, drawn from a normal distribution with a
scale set by the standard deviation of the continuum.  The reported
uncertainties are then drawn from the $84^{\rm th}$ and $16^{\rm th}$
percentiles of the resulting EW distribution.  We verified the overall
scale of our results by comparing the 21 stars that overlapped between
GALAH and Gaia-ESO.  17 of the 21 measurements agreed within
1-$\sigma$.  The remaining four were stars with $T_{\rm
eff}>6200\,{\rm K}$, for which \citet{randich_gaiaeso_2018} reported
Li EW detections ranging from $50$ to $75\,$m\AA, and for which we
found either null results or only slightly significant
($\approx$30m\AA) detections.

@@

%\subsection{Contamination fraction across the halo}
%\subsection{How did the halo form?}
%\subsection{Rotation in NGC\,2516}
%\subsection{The lithium-rotation correlation}
%\subsection{Age of NGC\,2516}

% \subsection{Contamination fraction across the halo}
% 
% The main aim this study is to verify whether stars
% in the halo of \cn\ are the same age as stars in the core, and
% secondarily to quantify the contamination rate.
% Figure~\ref{fig:physical_x_rotn}

@@

A separate hypothesis that we can test using the existing data is the
idea that rapid rotation can be explained through binarity.  Analyses
by {\it e.g.}, \citet{gillen_ngts_2020} and
\citet{simonian_rapid_2019} have reported that rapid rotation in young
clusters and in the field is correlated with binarity.  Related
effects could include tidal synchronization for the shortest period
binaries, or alternatively disk truncation that when combined with
magnetic locking {\bf (CITE)} could produce the rotation period
bimodality.

The lower panels of Figure~\ref{fig:rot} show our attempt at
identifying unresolved binaries in the \cn\ rotation sample.  Binaries
resolved by Gaia were already excluded in our initial definitions of
the samples.  In \cn\, we see that the fraction of stars showing signs
of binarity in the slow and fast rotation subsamples are 30\% (63/210)
and 56\% (40/71); the fast rotators show a preference for binarity.
This correlation is in line with the earlier findings noted above,
though we emphasize many rapid rotators in \cn\ (34\%) have no
evidence for being binary.

Comparing the explanations of disk-locking against tidal
synchronization, the population statistics do seem to strain the
possibility of tidal synchronization.  A quarter of the \cn\ members
are rapidly rotating.  In comparison, in the field half of Sun-like
stars are binaries, and $\approx$9\% have periods below 100\,days
\citep{raghavan_survey_2010}.  If we assume  that all binaries with
sub-100\,day periods become tidally synchronized, we can only explain
a rapid rotator occurrence rate of $\approx$5\%.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
carries over beyond just equivalent widths, and to the actual
abundance of lithium in the photosphere, A(Li).  To our knowledge, the
first demonstration of this was by \citet{soderblom_evolution_1993},
who calculated the relevant LTE curves of growth for the
Li\,6708\,\AA\ doublet, and used them to convert equivalent widths to
abundances.  Non-LTE corrections affect the abundances by $<0.5\,{\rm
dex}$ and do not change the overall relationships between Li abundance
and stellar temperature or rotation
\citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation since
\citet{soderblom_evolution_1993} have expanded these results in the
Pleiades, the Psc-Eri stream, and M\,35, among other clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
In \cn, \citet{jeffries_rotation_1998} previously observed a
similar trend that that seen in Figure~\ref{fig:lithiumrot} when
analyzing 24 stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
since these earlier studies.

% The fact that we see the same trend across the core and halo of \cn\ 
% {\it a)} supports the interpretation that the halo is coeval with the
% core, and {\it b)} suggests that the lithium-rotation correlation is
% not explained by environmental effects such as the density of the cluster
% during its embedded phase, but is instead tied to physical processes
% acting on all low-mass pre-main-sequence stars.

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?
{\bf FIXME: improve this paragraph... (e.g., see the introductory
material in Jeffries+21}
%FIXME IMPROVE THIS PARAGRAPH.
Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

\subsection{The Age of NGC\,2516}
\label{disc:absage}

\paragraph{Age from Gyrochronology}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrp<1.2$.  At redder colors ($1.2<\bpmrp<1.7$) the
dispersion in rotation periods increases.  The maximum rotation
periods seen in \cn\ across both sets $\mathcal{A}$ and $\mathcal{B}$
extend up to $\approx$11 days, rather than the $\approx$8.5day upper
limit seen in the Pleiades.  This is consistent with \cn\ being
slightly older than the Pleiades.

@@

\paragraph{Age from Lithium Depletion}

>>>>

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{The Halo of NGC\,2516}

@@

\title{
  Cluster Difference Imaging Photometric Survey. III.
  An Open Cluster Spanning 500\,Parsecs
}

@@

%
% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
%
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  It is
  important to verify that these ``halos'', ``tails'', and ``strings''
  are of similar ages and compositions as stars in the denser part of
  the cluster.  We present an analysis of NGC\,2516
  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
  and an apparent halo of stars spanning 500\,pc.  Combining
  photometry from Gaia, rotation periods from TESS, and lithium
  measurements from Gaia-ESO and GALAH, we find that the halo of
  NGC\,2516 is real and coeval with the cluster's core.  Two thirds of
  kinematically selected halo members out to 250\,pc from the cluster
  center have rotation periods consistent with a gyrochronological age
  of 150\,Myr.  A comparison sample of field stars shows no such
  trend.  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of
  similar clusters.  Broadly speaking, this work advances a paradigm
  in which the halos of open clusters are often more populous than
  their cores.  We highlight the implications for spectroscopic survey
  targeting, open cluster dispersal, and planet searches around young
  stars.
\end{abstract}

@@

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following intervals of right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample.

@@

An initial step in visualizing the reported structure and kinematics
of the candidate cluster members is to consider their Cartesian
coordinates (Figure~\ref{fig:XYZ}).  We computed these positions by
transforming from $(\alpha, \delta, \pi)$ to galactic ($X,Y,Z$)
assuming the \texttt{astropy v4.0} coordinate standard
\citep{astropy_2018}.  The direction of galactic rotation is
$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
median parameters for core members for which
\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.

@@

Another possible explanation for scatter in the halo's HR diagram
could be differential reddening across different sightlines.  The
reported halo spans 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun by up to 300\,pc
(Figure~\ref{fig:gaia6d}).  An HR diagram binned by galactic latitude
did show some minimal evidence for differential reddening, and so we
expect that both field star contamination and differential reddening
could play a role in the larger scatter of the halo relative to the
core.  A third possibility that we have not explored is whether the
binary fraction could also differ between the different regions of the
cluster.

@@

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).

Our goal is to determine whether the ages of the core and halo are
consistent.  The left panel of Figure~\ref{fig:hr} suggests that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The red-giant-branch
(RGB) stars are consistent only with the 178 Myr ($\log_{10} t/{\rm
yr} = 8.25$) model.  Based on the assumption that the width of the
MSTO can be attributed to binary stars, we are most interested in the
blue edge (blue stragglers are, however, a concern).  The blue edge
appears most compatible with the 178 Myr model.  These results are
consistent with the MSTO age range of 165 to 195\,Myr found by
\citet{cummings_2018}, and we refer to their work for the more precise
model-dependent comparison.  Based on Table~3 of
\citet{cummings_2018}, we therefore adopt a MSTO age for \cn\ of
$175\pm35$\,Myr, where our quoted uncertainty is a quadrature sum of
the statistical and systematic components from \cite{cummings_2018}.

@@

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in set $\mathcal{B}$.  The
``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrp<1.2$.  The implications of this comparison for the age of
\cn\ are discussed in Section~\ref{disc:absage}.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members that meet kinematic
  membership criteria (Subset~$\mathcal{A}$).  Stars with detected
  rotation periods are shown in blue;  halo members for which rotation
  periods should have been detectable, but were not detected, are
  shown in orange.  Stars in the core are shown as black points.
  Computing the 2D-tangential velocity requires correcting for
  projection effects (Appendix~\ref{app:vproj}).  In the {\it top
  right}, the histograms on bin versus 3D separation from the core and
  2D tangential velocity difference.  Bin widths are chosen to enforce
  the same number of stars in the denominator ($N=54$) per bin.  In the
  {\it bottom panels}, the innermost and outermost 100 stars in
  position and velocity space are displayed in the space of rotation
  periods and dereddened colors.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.9\textwidth]{f5a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with Gaia-ESO
  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Field stars (gray
  points) are Keck-HIRES measurements of stars with detected
  transiting planets from the Kepler mission
  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in violet
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections.
  {\bf LGB TODO: the increasing EWs for the Gaia-ESO M dwarfs must be erroneous. What's happening there?}
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f5b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
  GALAH surveys, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ (including
  non-detections) are shown with gray crosses.  The lithium-rotation
  correlation for the K-dwarfs has been observed in other young
  clusters, and is discussed in Section~\ref{subsec:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

@@

%
%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
%

We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.  We then measured the lithium equivalent widths using
the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
by first focusing in on a 15\AA\ window centered on the 6707.835\AA
lithium doublet.  We then continuum normalized the spectra using a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
to the continuum normalized spectra, considering only a $\pm 1\,$\AA
window centered on the Li doublet.  The equivalent widths were then
evaluated by numerically integrating the fitted model over the same
window.  Our approach therefore includes the Fe\,6707.44 blend in the
reported Li equivalent widths---which leads to systematic
overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
uncertainties on the EWs, we repeated the procedure twenty times, but
added noise to the spectra, drawn from a normal distribution with a
scale set by the standard deviation of the continuum.  The reported
uncertainties are then drawn from the $84^{\rm th}$ and $16^{\rm th}$
percentiles of the resulting EW distribution.  We verified the overall
scale of our results by comparing the 21 stars that overlapped between
GALAH and Gaia-ESO.  17 of the 21 measurements agreed within
1-$\sigma$.  The remaining four were stars with $T_{\rm
eff}>6200\,{\rm K}$, for which \citet{randich_gaiaeso_2018} reported
Li EW detections ranging from $50$ to $75\,$m\AA, and for which we
found either null results or only slightly significant
($\approx$30m\AA) detections.

@@

@@

A separate hypothesis that we can test using the existing data is the
idea that rapid rotation can be explained through binarity.  Analyses
by {\it e.g.}, \citet{gillen_ngts_2020} and
\citet{simonian_rapid_2019} have reported that rapid rotation in young
clusters and in the field is correlated with binarity.  Related
effects could include tidal synchronization for the shortest period
binaries, or alternatively disk truncation that when combined with
magnetic locking \citep[{\it
e.g.},][]{koenigl_disk_1991,long_locking_2005} could produce the
rotation period bimodality.

The lower panels of Figure~\ref{fig:rot} show our attempt at
identifying unresolved binaries in the \cn\ rotation sample.  Binaries
resolved by Gaia were already excluded in our initial definitions of
the samples.  In \cn\, we see that the fraction of stars showing signs
of binarity in the slow and fast rotation subsamples are 30\% (63/210)
and 56\% (40/71); the fast rotators show a preference for binarity.
This correlation is in line with the earlier findings noted above.
While many rapid rotators in \cn\ (34\%) have no evidence for being
binary, we emphasize that the cuts we used to select binaries (${\rm
RUWE}>1.2$; photometric excess exceeding 0.3\,mag) leave open a
significant fraction of parameter space.  A 0.3\,mag flux excess, for
instance, corresponds to mass ratios $M_2/M_1 \gtrsim 0.72$, assuming
the luminosity $L$ scales as $M^{3.5}$.

Comparing the explanations of disk-locking against tidal
synchronization, the population statistics seem to strain the
possibility of tidal synchronization.  A quarter of the \cn\ members
are rapidly rotating.  In comparison, in the field half of Sun-like
stars are binaries, and $\approx$9\% have periods below 100\,days
\citep{raghavan_survey_2010}.  If we assume that all binaries with
sub-100\,day periods become tidally synchronized, we can only explain
a rapid rotator occurrence rate of $\approx$5\%.  Elevated binarity
fractions in pre-MS stars likely do not change this picture \citep[see
Sec.~4.4 of][]{duchene_stellar_2013}.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) carries
over beyond just equivalent widths, and to the actual abundance of
lithium in the photosphere, A(Li).  To our knowledge, the first
demonstration of this was by \citet{soderblom_evolution_1993}, who
calculated the relevant LTE curves of growth for the Li\,6708\,\AA\
doublet, and used them to convert equivalent widths to abundances.
Non-LTE corrections affect the abundances by $<0.5\,{\rm dex}$ and do
not change the overall relationships between Li abundance and stellar
temperature or rotation \citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation have expanded these
results in the Pleiades, the Psc-Eri stream, and M\,35, among other
clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
In \cn, \citet{jeffries_rotation_1998} previously observed a similar
trend that that seen in Figure~\ref{fig:lithiumrot} when analyzing 24
stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
from this earlier analysis.

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?  This question has yet to be definitively
answered, but hints likely lie in shared correlations between lithium
abundance, rotation rate, radius inflation, internal mixing, and
magnetic field strength
\citep{chabrier_evolution_2007,somers_measurement_2017,jeffries_m35_li_2020}.
For instance, one interpretation of the lithium-rotation correlation
based on pure hydrodynamics is that the rapid rotation leads to less
efficient convective penetration, which then lowers the mixing
efficiency at the convective-radiative boundary
\citep{baraffe_lithium_2017}.  The magnetic field itself may even be
sufficient to inhibit the convection \citep{ventura_Li_B_1998}.
Alternatively, the star's structure could be at fault: rapid rotators
have the strongest magnetic fields and the most spotted surfaces.
These spots lower the photospheric flux, and drive the stellar radius
to expand while lowering the core temperatures, which in turn slows
the rate of lithium-destroying proton capture reactions
\citep{feiden_magnetic_2013,somers_rotation_2015}.

None of the internal processes noted above explain the initial
question of why some of the G and K dwarfs rotate rapidly to begin
with.  One clue noted above is that the rapid rotators tend to be in
photometric or astrometric binaries.  The majority of such binaries
have wide ($a\gtrsim200$\,au) or intermediate (0.5 - 100\,au)
separations \citep{raghavan_survey_2010}.

It is expected that the circumstellar disks in these binaries have
different properties than those of single stars, due to {\it e.g.} gap
formation in the circumbinary disk, disk truncation, and faster disk
clearing \citep{artymowicz_dynamics_1994,moe_impact_2020}.  It would
therefore be reasonable to assume that disk lifetimes in the binary
systems are shorter than in single star systems.
\citet{eggenberger_impact_2012} showed that longer disk lifetimes
should lead to an extended phase of disk-locking, which leads to a
greater amount of differential rotation generated in the radiative
zone, more efficient shear mixing, and ultimately a lower photospheric
lithium abundance.  This provides a plausible explanation for the
overall trend that the rapid rotators are often in unresolved
binaries, and that they are also lithium rich.  A careful
observational analysis including an assessment of the completeness to
different separations and mass ratios of binaries would be useful to
clarify whether or not all of the rapid rotators are in binaries.  If
not, the scenario presented above would be falsified.

\subsection{The age of NGC\,2516}
\label{disc:absage}

\paragraph{Age from Gyrochronology}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrp<1.2$.  At redder colors, $1.2<\bpmrp<1.7$, the
dispersion in rotation periods increases.  The maximum rotation
periods seen in \cn\ across both sets $\mathcal{A}$ and $\mathcal{B}$
extend up to $\approx$11 days, rather than the $\approx$8.5day upper
limit seen in the Pleiades.  This is consistent with \cn\ being
slightly older than the Pleiades.

@@

\paragraph{Age from Lithium Depletion}
The lithium depletion boundary for \cn\ has to our knowledge not been
measured.  Given the distance to the cluster, this is not surprising:
at $\sim$150\,Myr, the LDB is expected to occur at a stellar mass of
$\approx$0.088\,$M_\odot$ \citep{soderblom_ages_2014}.  This would
correspond to $\bpmrp\approx4.7$, well beyond the limits of available
cluster membership lists and spectroscopy.
